On existence of integral point sets and their diameter bounds \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2020-01-04 pps-linear-bound/pps-linear-lower-bound.tex
[J220] SciLexic to Acknowledgements
[J220] Review 2, medium remark 7

\begin{definition}
	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
	we define a $\rho(k,M_1 M_2)$-curve as the set of points $N$
	for which the equality $|N M_1| - |N M_2| = k$ holds.
\end{definition}
So, in the planar case a $\rho(k,M_1 M_2)$-curve
is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.

\begin{proposition}
	\label{obs:rho_curves}
	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
	the equality $|M_1 M_2| = |M_3 M_4|$ holds and line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and the $\rho(k,M_3 M_4)$-curve
	do not intersect.
\end{proposition}

\begin{lemma}
	\label{lem:2k-1_segments}
	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset l$ with endpoints in $M$,
	such that $|\Delta_i| = k$.
\end{lemma}
\begin{proof}
	Consider a point $N\in M \setminus l$.
	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
	otherwise the $\rho(n_i,\Delta_i)$-curve and the $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
	There can be only $2k-1$ distinct values for $n_i$,
	so there are at most $2k-1$ distinct segments $\Delta_i$.
\end{proof}

@@

\section{Acknowledgements}
The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
and Dr. A.S. Usachev for proofreading.
Also the author would like to owe a special thanks to the anonymous referees
for their helpful comments and suggestions on the paper.

>>>>

\begin{definition}
	For a line segment $M_1 M_2$ and an integer $k$, such that $-|M_1 M_2| < k < |M_1 M_2|$,
	we define a $\rho(k,M_1, M_2)$-curve as the set of points $N$
	for which the equality $|N M_1| - |N M_2| = k$ holds.
\end{definition}
So, in the planar case a $\rho(k,M_1, M_2)$-curve
is a branch of a hyperbola for $k\neq 0$ and the perpendicular bisector of line segment $M_1 M_2$ for $k=0$.

\begin{proposition}
	\label{obs:rho_curves}
	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
	the equality $|M_1 M_2| = |M_3 M_4| =h $ holds, line segments $M_1 M_2$ and $M_3 M_4$ do not coincide
	and vectors $\overrightarrow{M_1 M_2}$ and $\overrightarrow{M_3 M_4}$ are of equal direction,
	then for a fixed integer $k$, $|k| < h$, the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
	do not intersect.
\end{proposition}

\begin{proof}
	For $k=0$, the claim is obvious.
	Indeed, in this case the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve
	are distinct hyperplanes orthogonal to $l$ and thus do not intersect.

	For $k\neq 0$, suppose the contrary.
	Let $N$ be a point in the intersection.

	Consider a plane $\Pi \supset (l \cup \{N\})$
	with a Cartesian coordinate system such that $M_i = (x_i,0)$ for $1 \leq i \leq 4$.
	Let $R_1$ and $R_2$ be the projections
	of the $\rho(k,M_1, M_2)$-curve and the $\rho(k,M_3, M_4)$-curve onto $\Pi$.

	Then $R_1$ is a branch of a hyperbola defined by the equation
	\begin{equation}
		x- \frac{x_1+x_2}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
	\end{equation}
	and $R_2$ is a branch of a hyperbola defined by the equation
	\begin{equation}
		x- \frac{x_3+x_4}{2}=a(h,k)\frac{\sqrt{y^2+b^2(h,k)}}{b(h,k)}
		,
	\end{equation}
	where $a(h,k)$ and $b(h,k)$ are coefficients which depend on $h$ and $k$.
	For $N=(x_N, y_N) \in R_1 \cap R_2$ we have
	\begin{equation}
		\begin{cases}
			x_N- \dfrac{x_1+x_2}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
			\\[4pt]\\
			x_N- \dfrac{x_3+x_4}{2}=a(h,k)\dfrac{\sqrt{y_N^2+b^2(h,k)}}{b(h,k)}
		\end{cases}
	\end{equation}
	and, thus, $x_1+x_2 = x_3+x_4$.
	On the other hand, $x_2 - x_1 = x_4 - x_3 = h$,
	and we conclude that $x_1 = x_3$ as well as $x_2 = x_4$.
	So, line segments $M_1 M_2$ and $M_3 M_4$ coincide.
	The contradiction concludes the proof.
\end{proof}

\begin{lemma}
	\label{lem:2k-1_segments}
	Let $M \in \mathfrak{M}(2,\mathbb{N})$ and let $l$ be a straight line.
	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset l$ with endpoints in $M$,
	such that $|\Delta_i| = k$.
\end{lemma}
\begin{proof}
	Consider a point $N\in M \setminus l$.
	Let $\Delta_i = M_i^1 M_i^2$, and
	let vectors $\overrightarrow{M_i^1 M_i^2}$ and $\overrightarrow{M_j^1 M_j^2}$ be of equal direction for every $i$ and $j$.
	Then for each $i$ there is a $\rho(n_i, M_i^1, M_i^2)$-curve containing $N$.
	Due to Proposition~\ref{obs:rho_curves}, all $n_i$ are distinct;
	otherwise the $\rho(n_i,M_i^1, M_i^2)$-curve and the $\rho(n_i,M_j^1, M_j^2)$-curve, $j\neq i$, do not intersect.
	There can be only $2k-1$ distinct values for $n_i$,
	so there are at most $2k-1$ distinct segments $\Delta_i$.
\end{proof}

@@

\section{Acknowledgements}
The author thanks Dr. Prof. E.M. Semenov for the fruitful discussion and ideas,
A.S. Chervinskaia for the idea of using the term ``facher'' and proofreading,
Dr.~A.S.~Usachev for proofreading,
and E.A. Momot for the ``SciLexic'' project that helped the author with some word usage issues.
Also the author would like to owe a special thanks to the anonymous referees
for their helpful comments and suggestions on the paper.
